---
title: "BPWTP Monthly Report"
fontsize: 7pt
geometry: left = 1.5cm, right = 1.5cm, top = 2cm, bottom = 2cm
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
mainfont: Times New Roman
header-includes: "\\usepackage{float}\n\\usepackage{caption}\n\\usepackage{fancyhdr}\n\\pagestyle{fancy}\n\\fancyhead[L]{\\LARGE{BUFFALO
  POUND WATER TREATMENT PLANT}} \n\\setlength\\headheight{15pt}\n\\fancyhead[L]{\\large{Monthly
  Summary Report: Jul 2020}}\n\\fancyhead[L]{\\large{Monthly Summary Report: 2020}}\n"
editor_options:
  chunk_output_type: console
params:
  path_to_db_file: ""
  outdir: ""
  selected_year: NA
  selected_month: ""
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.pos = 'H', out.extra = "")
#code <- sapply(list.files(file.path(".", "code"), full.names = TRUE), source)
```

```{r}
library(bpwtpR)
library(pander)
library(tidyverse)
library(lubridate)
library(cowplot)
```


```{r update-data, echo = F, message = F, results = 'hide'}

## The complete labdat database file (labdat_file) from 1980 to 2019 will never be changed
## it will only be updated with the most contemporary data and saved as a new file

labdat <- read_db(params$path_to_db_file)
# To simplify column names while the data frame is being used. Blair wanted
  # columns to start with _tbl
colnames(labdat) <- str_remove(colnames(labdat), "tbl_")
```

```{r plot-all, eval = T}

plot_all_parameters(labdat, params$selected_year, 
                    parameter_group = "all",
                    station_list = "all",
                    params$outdir)

```



# Data Availability

Data in this report includes weekly sampling across the water treatment process at Buffalo Pound Water Treatment Plant (BPWTP) from `r min(year(labdat$date_ymd))` to `r max(year(labdat$date_ymd))`.

\newpage

# Operations

```{r operations_table}
summarize_monthly_table(df = labdat, 
            grouping = "operations", sampling_station = "Raw", 
            params$selected_year)
```

```{r operations_table_historical}

group_table_historical(df = labdat, params$selected_year,
                       grouping = "operations", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = T,
                       params$selected_month)

group_table_historical(df = labdat, params$selected_year,
                       grouping = "operations", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F,
                       params$selected_month)

```


\newpage

# Physical

```{r}
raw_odour <- parm_fig(df = labdat, 
                      params$selected_year, 
                      var = "Odour",
                      params$outdir)
print(raw_odour)
```

\newpage

## RawWater

```{r physical_table}
summarize_monthly_table(df = labdat, 
                        grouping = "physical", 
                        sampling_station = "Raw",
                        params$selected_year)
```

```{r physical_table_historical}
group_table_historical(df = labdat, 
                       params$selected_year, 
                       grouping = "physical", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = T, 
                       params$selected_month)
```


\newpage

# Major Constituents
## RawWater
```{r MC_table}
summarize_monthly_table(df = labdat, 
                        grouping = "majorConstituents", 
                        sampling_station = "Raw", 
                        params$selected_year)
```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year, 
                       grouping = "majorConstituents", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = T, 
                       params$selected_month)
```

\newpage

# Trace Constituents
## RawWater 
```{r TC_table}
summarize_monthly_table(df = labdat, 
                        grouping = "traceConstituents", 
                        sampling_station = "Raw", 
                        params$selected_year)
```

```{r }
group_table_historical(df = labdat, params$selected_year,
                       grouping = "traceConstituents",
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = T, 
                       params$selected_month)
```

## PreFM
```{r TCPreFM_table}
summarize_monthly_table(df = labdat, 
                        grouping = "traceConstituents", 
                        sampling_station = "PreFM", 
                        params$selected_year)
```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year, 
                       grouping = "traceConstituents", 
                       sampling_station = "PreFM", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = T,
                       params$selected_month)
```

\newpage

# THMs
```{r calc_thms}
thms <- summarize_THM(labdat = labdat)
```

```{r thm_fig, fig.height=9, fig.cap="Individual and total trihalomethane concentrations throughout the BPWTP treatment process. The maximum acceptable concentration for tTHM is 0.1 mg/L or 100 ug/L (effective 2016).", fig.scap="tTHM time series", fig.asp= 1}
tthm_fig(df = thms, 
         params$selected_year)
```

Tables are presented on the next page.

\newpage 

```{r thm_table-preFM}
summarize_monthly_table(df = labdat,  
                        grouping = "THM", 
                        sampling_station = "PreFM", 
                        params$selected_year)
```

```{r thm_table-channel}
summarize_monthly_table(df = labdat,
                        grouping = "THM", 
                        sampling_station = "Channel", 
                        params$selected_year)
```

```{r thm_table-pregac}
summarize_monthly_table(df = labdat, 
            grouping = "THM", sampling_station = "PreGAC", 
            params$selected_year)
```

```{r thm_table-clearwell}
summarize_monthly_table(df = labdat,
                        grouping = "THM", 
                        sampling_station = "Clearwell", 
                        params$selected_year)
```



\newpage

# Biological
## RawWater
```{r bio_table}
summarize_monthly_table(df = labdat,  
                        grouping = "biological", 
                        sampling_station = "Raw", 
                        params$selected_year)
```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year, 
                       grouping = "biological",
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = T, 
                       params$selected_month)
```

\newpage

# Bacteriological
## RawWater
```{r bac_table}
summarize_monthly_table(df = labdat, 
                        grouping = "bacteriological", 
                        sampling_station = "Raw", 
                        params$selected_year)
```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year, 
                       grouping = "bacteriological", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = T,  
                       params$selected_month)
```


\newpage

# Historical Data Summaries
## Operations

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "operations",
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F, 
                       params$selected_month)

```

\newpage

## Physical
```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "physical", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F,  
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "physical", 
                       sampling_station = "Clearwell", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F, 
                       params$selected_month)

```

\newpage

## Major Constituents
```{r }
group_table_historical(df = labdat,
                       params$selected_year,
                       grouping = "majorConstituents",
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F,
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "majorConstituents", 
                       sampling_station = "Clearwell", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9),
                       by_month = F, 
                       params$selected_month)

```

\newpage

## Trace Constituents
```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "traceConstituents", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9),
                       by_month = F,
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "traceConstituents",
                       sampling_station = "Clearwell", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F, 
                       params$selected_month)

```

\newpage

## tTHMs
```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "THM",
                       sampling_station = "PreFM",
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F, 
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat,
                       params$selected_year,
                       grouping = "THM",
                       sampling_station = "Channel", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9),
                       by_month = F, 
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "THM", 
                       sampling_station = "PreGAC", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9),
                       by_month = F, 
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "THM", 
                       sampling_station = "Clearwell",
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F, 
                       params$selected_month)

```

\newpage

## Biological
```{r }
group_table_historical(df = labdat,
                       params$selected_year,
                       grouping = "biological",
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F, 
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "biological", 
                       sampling_station = "Clearwell", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9),
                       by_month = F, 
                       params$selected_month)

```

\newpage

## Bacteriological
```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "bacteriological", 
                       sampling_station = "Raw", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9), 
                       by_month = F, 
                       params$selected_month)

```

```{r }
group_table_historical(df = labdat, 
                       params$selected_year,
                       grouping = "bacteriological", 
                       sampling_station = "Clearwell", 
                       percentiles = c(0.1, 0.25, 0.5, 0.75, 0.9),
                       by_month = F,
                       params$selected_month)

```

\listoffigures
\listoftables 
